---
title: "Mini Quiz"
author: "Ivan Lopez"
format: html
editor: visual
---

### QUIZ

- You will use the Zion National Park study data from `spDataLarge`: the park boundary (`zion.gpkg`), 30 random points inside the park (`zion_points`), and the Zion-area SRTM elevation raster (`srtm.tif`).

- The quiz below tracks the lecture’s main themes: simplification, buffers and clipping, affine transformations.

Use this setup chunk first.

```{r}

pacman::p_load(
  spDataLarge,
  sf,
  terra,
  tidyverse,
  rmapshaper
)

# Read the Zion study-area data
zion = read_sf(system.file("vector/zion.gpkg", package = "spDataLarge"))
data("zion_points", package = "spDataLarge")
srtm = rast(system.file("raster/srtm.tif", package = "spDataLarge"))

# Put everything in a projected CRS so distances are in meters
zion_utm = st_as_sf(project(vect(zion), "EPSG:32612"))
zion_points_utm = st_as_sf(project(vect(zion_points), "EPSG:32612"))
srtm_utm = project(srtm, "EPSG:32612")

```

## Question 1

Write code that simplifies the Zion National Park boundary using `ms_simplify()`, keeping 5% of the original vertices while preserving shape. Then plot the original and simplified boundaries side by side.

```{r}

zion_utm_simp = rmapshaper::ms_simplify(
  
  zion_utm,
                                          
  keep = 0.05, # Just keep 5% of the vertices!
                                          
  keep_shapes = TRUE  # keep the original adjacency
  
  )

plot(st_geometry(zion_utm), main="original")

plot(st_geometry(zion_utm_simp), main="congruent-simp")

packageVersion("Rcpp")
```

### Solution code

```{r}

zion_simpl = st_simplify(
  zion_utm,
  keep = 0.05,
  keep_shapes = TRUE
)

par(mfrow = c(1, 2))
plot(st_geometry(zion_utm), main = "Original")
plot(st_geometry(zion_simpl), main = "Incongruent")
par(mfrow = c(1, 1))
```

## Question 2

Write code that creates a 750-meter buffer around each point in `zion_points_utm`, clips those buffers to the Zion park boundary, and plots the result.

```{r}

# Encapsulate the Seine River in a 5km buffer

zion_buff_750m = st_buffer( #создания буферов (расширений вокруг геометрии).
  
  zion_points_utm,
  
  dist = 750
  
  )

clip=st_intersection(zion_buff_750m, zion_utm)

plot(st_geometry(zion_utm))
plot(st_geometry(clip), add=TRUE, main="750 m")
plot(st_geometry(zion_points_utm),add=T, main="15 km")
```

### Solution code

```{r}

zion_buffers = st_buffer(zion_points_utm, dist = 750)


zion_buffers_clip = st_intersection(zion_buffers, zion_utm)

plot(st_geometry(zion_utm), main = "750 m buffers clipped to Zion boundary")
plot(st_geometry(zion_buffers_clip), add = TRUE, col = "lightblue", border = "blue")
plot(st_geometry(zion_points_utm), add = TRUE, pch = 16, col = "red")


```

## Question 3

Using an affine transformation, write code that rotates the Zion park boundary 15 degrees clockwise around the centroid of the whole park. Plot the original and rotated boundaries together.

```{r}

rotation = function(a){
  
  r = a * pi / 180 #degrees to radians
  
  matrix(
    
    c(
      
      cos(r),
      
      sin(r),
      
      -sin(r),
      
      cos(r)),
    
    nrow = 2,
    
    ncol = 2)
  
} 

zion_sfc = st_geometry(zion_utm)

whole=st_union(zion_sfc) # объединяет части карты в один геом объект

whole_cent = st_centroid(whole)

zion_rotate = (zion_sfc - whole_cent) * rotation(15) + whole_cent 

plot(st_geometry(zion_utm), main = "rotation")

plot(st_geometry(zion_rotate), add=T, border="red")
```

### Solution code

```{r}

cent = st_coordinates(st_centroid(st_union(zion_utm)))

zion_centered = st_geometry(zion_utm) - cent

theta = -15 * pi / 180
rot_mat = matrix(
  c(cos(theta), -sin(theta),
    sin(theta),  cos(theta)),
  nrow = 2,
  byrow = TRUE
)

zion_rot = (zion_centered * rot_mat) + cent

plot(st_geometry(zion_utm), main = "Original and rotated Zion boundary", border = "black")
plot(zion_rot, add = TRUE, border = "red")
```
